### 2024-05-17 thesis/input_files/point.tex
May 17th 2024

This chapter covers the theory of the most widely used method for point prediction in the electricity forecasting literature. Besides discussing the theory underlying such models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of these models.

@@

\section{Autoregressive models}
% explain their theory
% and how the procedure how they are used
% IMPORTANT: load series is not a stationary series so before applying AR we have to perform stationary tests or differencing steps
Autoregressive models are a standard approach for modelling time series data.
Before introducing the popular model in electricity point forecasting, it is important to remember that autoregressive models assume the time series to be stationary. On the other hand, load and price have been observed to be a non stationary time series. 
Basically, stationarity means that the distribution of any subsequence of random variables of the stochatic process is invariant to shifts along the time dimension.
Therefore, checking stationarity is a crucial step in applying this class of models. %(unit root tests)
Furthermore, should the time series be non stationary, we have to convert it to a stationary one by a combination of either differencing, detrending and/or log transforming it.
\subsection{Autoregressive integrated moving average}
ARIMA(p,d,q) models are of the form
\begin{equation}
    \hat{y}_n=\sum\limits_{i=1}^{p}\psi_i y_{n-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{n-j}+\epsilon_n
\end{equation}
where the parameter p is the order of the autoregressive (AR) part, d is the degree of integrated differencing and q is the order of the moving average (MA) model. $\epsilon_{n-j}$ are the observed errors in the past while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
\\
ARIMA model are compactly written as
\begin{equation}
    \psi(B)\nabla^d y_t=\theta(B)\epsilon_t
\end{equation}
where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\psi(B)$ is shorthand for $\psi(B)=1-\psi_1B-\dots-\psi_p B^p$.
Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
\\
The most popular procedure for estimating this class of models is the Box-Jenkins method \cite{box2015time}.
The order p is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
In the same way, the order q is estimated by considering the sample autocorrelation function of the observed time series. Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria. 
% Notice, models are obtained through the maximum likelihood estimation framework.

\subsection{Autoregressive integrated moving average with exogenous variables}
ARIMAX model adds explanatory variables to our ARIMA, hence, we have
\begin{equation}
    \hat{y}_n=\sum\limits_{k=1}^{h} \mu_i X_{n-k}+\sum\limits_{i=1}^{p}\psi_i y_{n-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{n-j}+ \epsilon_{n}
\end{equation}
\subsection{Seasonal autoregressive integrated moving average and Seasonal autoregressive integrated moving average with exogenous variables}
SARIMA and SARIMAX incorporate seasonalities into our time series models. The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$. The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal patter, where $m$ indicates the number of seasonalities of our time series.
For instance, in modelling series of hourly data with daily periodicity, we would set m=24.
Writing SARIMA in compact notation we have
\\
\begin{equation}
    \psi(B)\Psi(B^s)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^s)\epsilon_t
\end{equation}
where $\Psi$ and $\Theta$ account for the $P$ and $Q$ of the seasonal patter $m$ respectively.
\section{Generalized additive model}
% in standard regression the predicted y are normal since error are normal. But if we want to model say a y bernoulli which is the rate of success we can link the domain of bernoulli to that of gaussian with a link function
Generalized additive models (GAMs) take the form
\begin{equation}
    g[\mathbb{E}(Y_i)]=\alpha+f_1(X_1)+\dots+f_p(X_p)
\end{equation}
$f_i$ are smooth functions which are fitted using cubic smoothing splines.
g is called link function and characterizes the specific GAM model \cite{hastie2017generalized}. Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.

%prophet meta
Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}, developed by Meta. This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonality. Thus, the reason we choose it as a valid benchmark to compare against.
\section{K-nearest neighbours regression}
K-nearest neighbours regression forecasts by averaging the most similar k instances in the training set \cite{macqueen1967some}. K is the algorithm hyperparameter, it stands for the number of neighbours; a small k leads to overfitting while a big k leads to underfitting. 
Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see \ref{appendix:normalization}.
This algorithm is made up of two decisions: first, the choice of the metric and second, the method for combining targets.

@@

Developed at the AT\&T Bell Laboratories by Vapnik et al. \cite{cortes1995support} \cite{vapnik1997support}, support vector machines SVMs are one of the most popular techniques within the field of statistical learning.

@@

Put differently, we would like a model to keep error less than the $\epsilon$ threshold,  

@@

In standard SVR we have

@@

    f(x)=\langle w,x \rangle +b \ \textrm{with} \ w \in \mathcal{X}, b \in \mathbb{R}

@@

Where $\mathcal{X}$ denotes the space of the input patters $x_i$.

@@

The C constant trades off between $\epsilon$ deviation tolerance and flatness of the function f. A bigger C gives more importance to error minimization while as C gets smaller, flattness gains importance.

@@

        \max_{\alpha_i, \alpha_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\alpha_i-\alpha_i^*)(\alpha_j-\alpha_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)(y_i-\epsilon)

@@

        s.t. \quad& \sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)=0

@@

        \quad& \alpha_i, \alpha_i^* \in [0, C]

@@

Rearranging the gradient of the lagrangian with respect to $w$, we obtain the so called support vector expansion.

@@

    w=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)x_i

@@

Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those point $i$ for which $(\alpha_i-\alpha_i^*)\neq 0$

@@

Furthermore, equation \ref{eq:support vector expansion} implies

@@

    f(x)=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)\langle x_i, x\rangle +b

@@

Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. Particulalry, the condition that the product between the constraints and the dual variable has to vanish. These implies that the lagrange multipliers $\alpha_i, \alpha_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube. Consequently for any of these data points, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ holds.

@@

To get an idea of how SVR works, see figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.

@@

    \caption{support vector regression}

@@

\section{Kernel methods}
\subsection{Kernel ridge regression}
In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate $\hat{f}$ by solving the optmisation problem
\begin{equation}\label{eq:kernel ridge}
    \hat{f}=\argmin{f\in \mathcal{H}}\frac{1}{2}\sum\limits_{i=1}^n(y_i-f(x_i))^2+\frac{\lambda}{2}\|f\|_{\mathcal{H}}^2
\end{equation}
Before proceeding, we need to introduce the representer theorem.
\begin{theorem}
    Let $k:\mathcal{X} \times \mathcal{X} \mapsto R$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
    a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
    Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$. Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
\end{theorem}

\begin{proof}
    By orthogonal projection, any $f \in \mathcal{H}$ can be decomposed as the sum of two functions, with the first lying in the $\textrm{span}\{\phi(x_1),\dots, \phi(x_n)\}$ and the second staying in the orthogonal complement of the former. Letting $\nu$ denote a function from the orthogonal complement, we have
    \begin{equation}
        f=\sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu
    \end{equation}
    Applying $f$ to any point, by the reproducing property we have
    \begin{equation}
        \begin{aligned}
            f(x_j)=&\langle \phi(x_j), \sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu\rangle
            \\
            =&\sum\limits_{i=1}^{n}\alpha_i\langle \phi(x_j), \phi(x_i)\rangle
        \end{aligned}
    \end{equation}
    This means that setting $\nu=0$ does not affect the evaluation of $f$.
    Next consider the regularization term; notice, the functional $g$ is defined to be strictly monotonic in order to penalize more complex functions $f$.
    \begin{equation}
        \begin{aligned}
            g(\|f\|)=&g\left(\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}\right)
            \\
            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}^2}\right)
            \\
            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\left\|v\right\|_\mathcal{H}^2+2\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \rangle_{\mathcal{H}}}\right)
            \\
            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2}\right)
            \\
            \geq &g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2}\right)
        \end{aligned}
    \end{equation}
    Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected. Thus it follows that any minimizer of the risk functional of equation \ref{eq:kernel ridge} must be of the following form
    \begin{equation}
        f^*(\cdot)=\sum\limits_{i=1}^n \alpha_i \phi(x_i)=\sum\limits_{i=1}^n \alpha_i k(\cdot, x_i)
    \end{equation}
\end{proof}

Employing the representer theorem, we can rewrite \ref{eq:kernel ridge} in matrix notation as
\begin{equation}
    \begin{aligned}
    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}^2
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\lambda}{2} \alpha^\intercal K \alpha
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\lambda}{2} \alpha^\intercal K \alpha
\end{aligned}
\end{equation}

Setting the gradient to zero
\begin{equation}
    -Ky+K^2\alpha+\lambda K \alpha\overset{!}{=}0
\end{equation}

We have then, that $\hat{\alpha}=(K+\lambda I)^{-1}y$. It follows that the solution takes the form
\begin{equation}
    \hat{f}(\cdot)=\hat{\alpha}K(\cdot,:)
\end{equation}

\subsection{Kernel support vector regression}
Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean dot product of data $x$ with the dot product in the higher feature space $\mathcal{F}$. Doing so, the optimisation problem can be restated as
\begin{equation}
    \begin{aligned}
        \max_{\alpha_i, \alpha_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\alpha_i-\alpha_i^*)(\alpha_i-\alpha_i^*)k(x_i, x_j)+\sum\limits_{i=1}^n(y_i-\epsilon)(\alpha_i-\alpha_i^*)
        \\
        s.t. \quad& \sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)=0
        \\
        \quad& \alpha_i, \alpha_i^* \in [0,C]
    \end{aligned}
\end{equation}
Our regressor $f$ is then given by
\begin{equation}
    \begin{aligned}
        f(x)=\sum\limits_{i=1}^n(\alpha_i-\alpha_i^*)k(x_i, x)+b
    \end{aligned}
\end{equation}
Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel support vector regression we seek for the flattest function in the feature space not the input space.
See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former, we have used a polynomial kernel while in the latter the standard radial basis function kernel.
\begin{figure}
    \includegraphics[width=\textwidth]{images/svr2.png}
    \caption{Polynomial support vector regression}
    \label{fig:svr2}
\end{figure}

>>>>

This chapter covers the theory of the most widely used methods for point prediction in the electricity forecasting literature. Besides discussing the theory underlying such models, this and the following chapter include also a couple of worked examples in order to get acquainted with the practical applications of these models.

@@

\section{Autoregressive models}
% explain their theory
% and how the procedure how they are used
% IMPORTANT: load series is not a stationary series so before applying AR we have to perform stationary tests or differencing steps
Autoregressive models are a standard approach for modelling time series data.
Before introducing this popular model in electricity point forecasting, it is important to remember that autoregressive models assume the time series to be stationary. On the other hand, load and price have been observed to be a non stationary time series. 
In a nutshell, stationarity means that the distribution of any subsequence of random variables of the stochastic process is invariant to shifts along the time dimension.
Therefore, checking stationarity is a crucial step in applying this class of models. %(unit root tests)
Furthermore, should the time series be non stationary, we can convert it to a stationary one by a combination of either differencing, detrending and/or log transforming it.
\subsection{Autoregressive integrated moving average}
ARIMA(p,d,q) models are of the form
\begin{equation}
    \hat{y}_n=\sum\limits_{i=1}^{p}\psi_i y_{n-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{n-j}+\epsilon_n
\end{equation}
where the parameter p is the order of the autoregressive (AR) part, d is the degree of integrated differencing and q is the order of the moving average (MA) model. $\epsilon_{n-j}$ are the observed errors in the past while the differenced term $y_n^{(d)}$ is defined recursively as $y_n^{(d)}=y_n^{(d-1)}-y_{n-1}^{(d-1)}$.
\\
ARIMA model are compactly written as
\begin{equation}
    \psi(B)\nabla^d y_t=\theta(B)\epsilon_t
\end{equation}
where $B$ is the backward shift operator, $B^h y_t=y_{t-h}$ and $\psi(B)$ is shorthand for $\psi(B)=1-\psi_1B-\dots-\psi_p B^p$.
Similarly, $\theta(B)=1+\theta_1B+\dots+\theta_q B^q$ and $\nabla_h$ is the lag h differencing operator $\nabla_h y_t=(1-B^h)y_t=y_t-y_{t-h}$
\\
The most popular procedure for estimating this class of models is the Box-Jenkins method \cite{box2015time}.
The order p is identified by taking the lags at which the sample partial autocorrelation function (PACF) falls outside its 95\% confidence interval.
In the same way, the order q is estimated by considering the sample autocorrelation function (ACF) of the observed time series. Finally, we compare ARIMA models with different p,d,q (in a neighbourhood of the identified orders) and choose the ones minimising either AIC or BIC criteria. 
% Notice, models are obtained through the maximum likelihood estimation framework.

\subsection{Autoregressive integrated moving average with exogenous variables}
ARIMAX model adds explanatory variables to our ARIMA, hence, we have
\begin{equation}
    \hat{y}_n=\sum\limits_{k=1}^{h} \mu_i X_{n-k}+\sum\limits_{i=1}^{p}\psi_i y_{n-i}^{(d)}+\sum\limits_{j=1}^{q}\theta_j \epsilon_{n-j}+ \epsilon_{n}
\end{equation}
\subsection{Seasonal autoregressive integrated moving average and seasonal autoregressive integrated moving average with exogenous variables}
SARIMA and SARIMAX incorporate seasonalities into our time series models. The notation for these models is ARIMA$(p,d,q)(P,D,Q)_m$. The second term $(P,D,Q)_m$ represents the autoregressive structure of the seasonal pattern, where $m$ indicates the number of seasonalities of our time series.
For instance, in modelling series of hourly data with daily periodicity, we would set m=24.
Writing SARIMA in compact notation we have
\\
\begin{equation}
    \psi(B)\Psi(B^P)\nabla^d\nabla_m^Dy_t=\theta(B)\Theta(B^Q)\epsilon_t
\end{equation}
where $\Psi$ and $\Theta$ account for the $P$ and $Q$ of the seasonal pattern $m$ respectively.
\section{Generalized additive model}
% in standard regression the predicted y are normal since error are normal. But if we want to model say a y bernoulli which is the rate of success we can link the domain of bernoulli to that of gaussian with a link function
Generalized additive models (GAMs) take the form
\begin{equation}
    g[\mathbb{E}(Y_i)]=\alpha+f_1(X_1)+\dots+f_p(X_p)
\end{equation}
$f_i$ are smooth functions which are fitted using cubic smoothing splines.
g is called link function and characterizes the specific GAM model \cite{hastie2017generalized}. Approximations for $f_i$ are obtained through an iterative procedure, the backfitting algorithm.

%prophet meta
Particularly successful in time series forecasting is the Prophet model \cite{taylor2018forecasting}, developed by Meta. This additive model is capable of handling non linear trends, holiday effects and yearly weekly and daily seasonalities. Thus, the reason why we choose it as a valid benchmark to compare against.
\section{K-nearest neighbours regression}
K-nearest neighbours regression forecasts by averaging the most similar k instances in the training set \cite{macqueen1967some}. K is the algorithm hyperparameter, it stands for the number of neighbours; a small k leads to overfitting while a big k leads to underfitting. 
Since it is a metric based algorithm, it is important to normalize data in order to give equal weights to features with different scales, see \ref{appendix:normalization}.
This algorithm is made up of two decisions: first, the choice of the metric and second, the method for combining targets.

@@

Developed at the AT\&T Bell Laboratories by Vapnik et al. \cite{cortes1995support} \cite{vapnik1997support}, support vector machines (SVMs) are one of the most popular techniques within the field of statistical learning.

@@

Put differently, we would like a model to keep error less than the $\epsilon$ threshold.  

@@

In standard support vector regression (SVR) we have

@@

    f(x)=\langle w,x \rangle_{\mathbb{R}^n} +b \ \textrm{with} \ w \in \mathcal{X} \subseteq \mathbb{R}^n, b \in \mathbb{R}

@@

Where $\mathcal{X}$ denotes the space of the input data $x_i$.

@@

The $C$ constant trades off between $\epsilon$ deviation tolerance and flatness of the function $f$. A bigger $C$ gives more importance to error minimization while as $C$ gets smaller, flattness gains importance.

@@

        \max_{\lambda_i, \lambda_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\lambda_i-\lambda_i^*)(\lambda_j-\lambda_j^*)\langle x_i,x_j\rangle +\sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)(y_i-\epsilon)

@@

        s.t. \quad& \sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)=0

@@

        \quad& \lambda_i, \lambda_i^* \in [0, C]

@@

Rearranging the gradient of the primal lagrangian with respect to $w$, we obtain the so called support vector expansion.

@@

    w=\sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)x_i

@@

Notice, the complexity of the function representation is independent of the feature space dimensionality, but depends only on the number of support vectors; that is those point $i$ for which $(\lambda_i-\lambda_i^*)\neq 0$

@@

That is because, equation \ref{eq:support vector expansion} implies

@@

    f(x)=\sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)\langle x_i, x\rangle +b

@@

Employing the Karush-Kuhn-Tucker conditions, $b$ can be retrieved easily. Particulalry, the condition that the product between the constraints and the dual variable has to vanish.
Consequently for any of the data points with associated $\lambda_i, \lambda^*_i \in (0,C)$, the equality $b=y_i-\langle w, x_i-\epsilon\rangle$ has to hold.

@@

Furthermore, this implies that the Lagrange multipliers $\lambda_i, \lambda_i^*$ may be nonzero only for the samples inside the epsilon insensitive tube, that is only those points may be support vectors. 
To get an idea of how the algorithm works, see figure \ref{fig:svr1} for how a support vector regression handles a sinusoidal function with noise.

@@

    \caption{Linear support vector regression}

@@

\section{Kernel methods}
\subsection{Kernel ridge regression}
In the setting of kernel ridge regression, we aim at estimating $f$ such that $y=f(x)+\epsilon$. Given a RKHS $\mathcal{H}$, we can estimate $\hat{f}$ by solving the optmisation problem
\begin{equation}\label{eq:kernel ridge}
    \hat{f}=\argmin{f\in \mathcal{H}}\frac{1}{2}\sum\limits_{i=1}^n(y_i-f(x_i))^2+\frac{\tau}{2}\|f\|_{\mathcal{H}}^2
\end{equation}
Before proceeding, we need to introduce the representer theorem.
\begin{theorem}
    Let $k:\mathcal{X} \times \mathcal{X} \mapsto \mathbb{R}$ on a non empty set $\mathcal{X}$ with corresponding RKHS $\mathcal{H}$,
    a strictly increasing real valued function $g:[0,\infty) \to \mathbb{R}$, and the loss function $E:(\mathcal{X}\times \mathbb{R}^2)^n \to \mathbb{R} \cup \{\infty\}$.
    Consider the regularized risk functional $Risk[f]:f \mapsto  E\left((x_{1},y_{1},f(x_{1})),\ldots, (x_{n},y_{n},f(x_{n}))\right)+g\left(\| f\| \right)$. Then any minimizer of the empirical risk $Risk[f]$ can be represented as $ f^{*}(\cdot )=\sum\limits_{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})$
\end{theorem}

\begin{proof}
    By orthogonal projection, any $f \in \mathcal{H}$ can be decomposed as the sum of two functions, with the first lying in the $\textrm{span}\{\phi(x_1),\dots, \phi(x_n)\}$ and the second staying in the orthogonal complement of the former. Letting $\nu$ denote a function from the orthogonal complement, we have
    \begin{equation}
        f=\sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu
    \end{equation}
    Applying $f$ to any point, by the reproducing property we have
    \begin{equation}
        \begin{aligned}
            f(x_j)=&\langle \phi(x_j), \sum\limits_{i=1}^{n}\alpha_i \phi(x_i)+\nu\rangle_{\mathcal{H}}
            \\
            =&\sum\limits_{i=1}^{n}\alpha_i\langle \phi(x_j), \phi(x_i)\rangle_{\mathcal{H}}
        \end{aligned}
    \end{equation}
    This means that setting $\nu=0$ does not affect the evaluation of $f$.
    Next consider the regularization term; notice, the functional $g$ is defined to be strictly monotonic in order to penalize more complex functions $f$.
    \begin{equation}
        \begin{aligned}
            g(\|f\|_{\mathcal{H}})=&g\left(\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}\right)
            \\
            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)+v\right\|_\mathcal{H}^2}\right)
            \\
            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\left\|v\right\|_\mathcal{H}^2+2\left\langle\sum\limits_{i=1}^n \alpha_i \phi(x_i),\nu \right\rangle_{\mathcal{H}}}\right)
            \\
            =&g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2+\|v\|_\mathcal{H}^2}\right)
            \\
            \geq &g\left(\sqrt{\left\|\sum\limits_{i=1}^n \alpha_i \phi(x_i)\right\|_{\mathcal{H}}^2}\right)
        \end{aligned}
    \end{equation}
    Hence, we can conclude that setting $\nu=0$ decreases the second term while the first term is not affected. Thus it follows that any minimizer of the risk functional of equation \ref{eq:kernel ridge} must be of the following form
    \begin{equation}
        f^*(\cdot)=\sum\limits_{i=1}^n \alpha_i \phi(x_i)=\sum\limits_{i=1}^n \alpha_i k(\cdot, x_i)
    \end{equation}
\end{proof}

Employing the representer theorem, we can rewrite \ref{eq:kernel ridge} in matrix notation as
\begin{equation}
    \begin{aligned}
    \hat{\alpha}=&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2}\left\|\sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i})\right\|_{\mathcal{H}}^2
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \left\langle \sum _{i=1}^{n}\alpha _{i}k(\cdot ,x_{i}),\sum _{j=1}^{n}\alpha _{j}k(\cdot ,x_{j})\right\rangle_{\mathcal{H}}^2
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}\langle k(\cdot ,x_{i}),k(\cdot ,x_{j})\rangle_{\mathcal{H}}^2
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \sum _{i,j=1}^{n}\alpha _{i}\alpha _{j}K_{ij}
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}\|y-K\alpha\|_2^2+\frac{\tau}{2} \alpha^\intercal K \alpha
    \\
    =&\argmin{\alpha \in \mathbb{R}^n}\frac{1}{2}(y-K\alpha)^\intercal (y-K\alpha)+\frac{\tau}{2} \alpha^\intercal K \alpha
\end{aligned}
\end{equation}

Setting the gradient to zero
\begin{equation}
    -Ky+K^2\alpha+\tau K \alpha\overset{!}{=}0
\end{equation}

We have then, that $\hat{\alpha}=(K+\tau I)^{-1}y$. It follows that the solution takes the form
\begin{equation}
    \hat{f}(\cdot)=\hat{\alpha}K(\cdot,:)
\end{equation}

\subsection{Kernel support vector regression}
Support vector regression (SVR) can be kernelized by swapping the $\mathbb{R}^n$ euclidean dot product of data $x$ with the dot product in the higher feature space $\mathcal{H}$. Doing so, the optimisation problem can be restated as
\begin{equation}
    \begin{aligned}
        \max_{\lambda_i, \lambda_i^*} \quad& -\frac{1}{2}\sum\limits_{i,j=1}^n(\lambda_i-\lambda_i^*)(\lambda_i-\lambda_i^*)k(x_i, x_j)+\sum\limits_{i=1}^n(y_i-\epsilon)(\lambda_i-\lambda_i^*)
        \\
        s.t. \quad& \sum\limits_{i=1}^n(\lambda_i-\lambda_i^*)=0
        \\
        \quad& \lambda_i, \lambda_i^* \in [0,C]
    \end{aligned}
\end{equation}
Letting $\alpha_i=\lambda_i - \lambda_i^*$, our regressor $f$ is then given by
\begin{equation}
    \begin{aligned}
        f(x)=\sum\limits_{i=1}^n \alpha_i k(x_i, x)+b
    \end{aligned}
\end{equation}
Notice, in this setting $w$ is no longer given explicitly. Additionally, with kernel support vector regression we seek for the flattest function in the feature space not the input space.
See figure \ref{fig:svr2} and figure \ref{fig:svr3} for two examples. In the former, we have used a polynomial kernel while in the latter the standard radial basis function kernel.
\begin{figure}
    \includegraphics[width=\textwidth]{images/svr2.png}
    \caption{Polynomial support vector regression}
    \label{fig:svr2}
\end{figure}
